---
title: "Machine Learning Hello World"
output: html_notebook
---


load caret
```{r}
library(caret)
library(tidyverse)
```
load data
```{r}
dataset <- iris
```

We now split the dataset into data used for training (80%) and for testing (20%)
```{r}
# create a list of 80% of the rows in the original dataset we can use for training
validation_index <- createDataPartition(dataset$Species, p=0.80, list=FALSE)
# select 20% of the data for validation
validation <- dataset[-validation_index]
# use the remaining 80% of data to training and testing the models
dataset_80 <- dataset[validation_index,]
```
now we inspect the data
```{r}
head(dataset)
dim(dataset)
dim(dataset_80)
```
dataset has 150 instances and 5 attributes
dataset_80 has 120 instances and 5 attributes

We now investigate the types for each attribute
```{r}
sapply(dataset, class)
```
Notice column `Species` is of type factor. What levels are there for `Species`?
```{r}
levels(dataset$Species)
```
There are 3 possible species, hence we are dealing with a multinomial classification problem.

Now, how many representatives of each species are there?
```{r}
dataset %>% group_by(Species) %>% summarise(n())
```
The original dataset contains 50 representatives of each species
```{r}
dataset_80 %>% group_by(Species) %>% summarise(n())
```
The training dataset contains 40 representatives of each species

Now let's visualise the training dataset using box and whiskers diagrams
```{r}
# split dependant and independant variables
x <- dataset_80[,1:4]
# select only name of species
y <- dataset_80[,5]
```

```{r}
# boxplot for each attribute on one image
par(mfrow=c(1,4))
  for(i in 1:4) {
  boxplot(x[,i], main=names(iris)[i])
}
```








